---
title: "Summer Undergraduate Research Project"
authors: "Matina Lampsas & Manning Kalish"
output:
  html_document:
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
    code-fold: True
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(readxl)
```

```{r}
adjusted_data <- read.csv(here::here("data",
                     "adjusted_data_new.csv")
                     )
combined_data <- read.csv(here::here("data",
                     "house_data.csv")
                     )

```

```{r}
# make dummy vairable for SLO 

adjusted_data <- adjusted_data %>% 
  mutate(SLO = if_else(City == "SLO", 1,0))


slo_zip <- c("93401", "93403", "93405", "93406", "93408", "93410")

combined_data <- combined_data %>% 
  mutate(SLO = if_else(ZIP.OR.POSTAL.CODE %in% slo_zip, 1,0))




```
Atascadero: historic avg high: 80.6, avg.low: 54.4
Nipomo: 70.4 - high, 50.6 low
Paso Robles: high - 78.3, low - 54.2
San Luis Obispo: 64.9 high , 43.9 low
Arroyo Grande: 64.5 - high, 43.6 low
Pismo Beach	:high 80.6, 54.5 
Oceano: 68.2 high, 46.5 low
Grover Beach: 66.3 high, 45.4 low
Morro Bay: 78.2 high, 53.8 low
Avila Beach: 78.3 high, 54.2 low
Los Osos 66.2 high , 45.3 low
San Miguel: 65.9 high, 44.8 low
Santa Margarita	80.6 high, 54.4 low
Templeton: 80.6 high, 54.3 low
Cambria: 80.6, 54.4
Creston: 82.07, 64.74
Cayucos: 73.8, 50.8
Shandon: 78.1, 53.7
Bradley: 79.9, 53.5
Santa Maria: 73.4, 55.1
California Valley	78.5, 61.4
San Simeon: 68.4, 46.5

```{r}
# Create the data frame to add temperatures 
temp_data <- data.frame(
  CITY = c(
    "Atascadero", "Nipomo", "Paso Robles", "San Luis Obispo", "Arroyo Grande",
    "Pismo Beach", "Oceano", "Grover Beach", "Morro Bay", "Avila Beach",
    "Los Osos", "San Miguel", "Santa Margarita", "Templeton", "Cambria",
    "Creston", "Cayucos", "Shandon", "Bradley", "Santa Maria",
    "California Valley", "San Simeon"
  ),
  Avg_High_Temp = c(
    80.6, 70.4, 78.3, 64.9, 64.5, 80.6, 68.2, 66.3, 78.2, 78.3,
    66.2, 65.9, 80.6, 80.6, 80.6, 82.07, 73.8, 78.1, 79.9, 73.4,
    78.5, 68.4
  ),
  Avg_Low_Temp = c(
    54.4, 50.6, 54.2, 43.9, 43.6, 54.5, 46.5, 45.4, 53.8, 54.2,
    45.3, 44.8, 54.4, 54.3, 54.4, 64.74, 50.8, 53.7, 53.5, 55.1,
    61.4, 46.5
  )
)

```


```{r}
# rename  the city variable and then add the columns for the temps
adjusted_data <-adjusted_data %>%
  rename(CITY = City)
adjusted_data <- left_join(adjusted_data, temp_data, by = 'CITY')
combined_data <- left_join(combined_data, temp_data, by = 'CITY')
```


## Graphs - Week of 7/15

```{r}
city_filter <- c("Paso Robles", "San Luis Obispo", "Atascadero", "Arroyo Grande", "Nipomo", "Cambria")
combined_data %>% 
  filter(CITY %in% city_filter) %>% 
  ggplot(mapping = aes(x = PRICE)) +
  geom_histogram(binwidth = 10000, fill = "blue", color = "black", alpha = 0.7) +  # Adjust binwidth as necessary
  facet_wrap(~ CITY, scales = "free")+
  scale_x_continuous()
```
```{r}
city_filter <- c("Paso Robles", "San Luis Obispo", "Atascadero", "Arroyo Grande", "Nipomo", "Cambria")
combined_data %>% 
  filter(CITY %in% city_filter) %>% 
  summary()
```

```{r}
adjusted_data %>%
  count(City) %>%
  arrange(desc(n)) %>%
  mutate(City = factor(City, levels = City)) %>%
  ggplot(mapping = aes(x = City, y = n, fill = City)) +
  geom_bar(stat = "identity") +
  labs(x = "City", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

```{r}
adjusted_data %>% 
  select(Adjusted_Price) %>% 
  summary()
```
```{r}
adjusted_data %>%
  ggplot(mapping = aes(x = Adjusted_Price)) +
  geom_histogram(bins = 100, fill = "skyblue", color = "black") +
  labs(x = "Adjusted Price", y = "Frequency", title = "Histogram of Adjusted Prices")
```

```{r}
qnt <- adjusted_data %>%
  summarize(
    Q1 = quantile(Adjusted_Price, 0.25, na.rm = TRUE),
    Q3 = quantile(Adjusted_Price, 0.75, na.rm = TRUE)
  )

# Calculate the IQR
qnt <- qnt %>%
  mutate(
    IQR = Q3 - Q1,
    lower_limit = Q1 - 1.5 * IQR,
    upper_limit = Q3 + 1.5 * IQR
  )

# Filter outliers
adjusted_data <- adjusted_data %>%
  filter(Adjusted_Price >= qnt$lower_limit, Adjusted_Price <= qnt$upper_limit)

adjusted_data %>% 
  select(Adjusted_Price) %>% 
  summary()


adjusted_data %>%
  ggplot(mapping = aes(x = Adjusted_Price)) +
  geom_histogram(bins = 50, fill = "skyblue", color = "black") +
  labs(x = "Adjusted Price", y = "Frequency", title = "Histogram of Adjusted Prices")
```
```{r}
city_filter <- c("Paso Robles", "San Luis Obispo", "Atascadero", "Arroyo Grande", "Nipomo", "Cambria")

# # Define the mapping dictionary
# city_mapping <- list(
#   'TTON' = 'Templeton',
#   'GRVC' = 'Grover Beach',
#   'SLO' = 'San Luis Obispo',
#   'PSOR' = 'Paso Robles',
#   'NPMO' = 'Nipomo',
#   'ATSC' = 'Atascadero',
#   'UNIC' = 'Unincorporated',
#   'ARRG' = 'Arroyo Grande',
#   'PSMO' = 'Pismo Beach',
#   'AVIL' = 'Avila Beach',
#   'CAMB' = 'Cambria',
#   'SSIM' = 'San Simeon',
#   'CAYU' = 'Cayucos',
#   'CRST' = 'Creston',
#   'MRBY' = 'Morro Bay',
#   'OCNO' = 'Oceano',
#   'OSOS' = 'Los Osos',
#   'BDLY' = 'Bradley',
#   'SHDN' = 'Shandon',
#   'SMIG' = 'San Miguel',
#   'SMRG' = 'Santa Margarita',
#   'SMIA' = 'Santa Maria',
#   'LKNA' = 'Lake Nacimiento',
#   'CVLL' = 'Creston Village'
# )
# 
# # Replace city codes with actual names
# adjusted_data <- adjusted_data %>% 
#   mutate(City = recode(City, !!!city_mapping))


adjusted_data %>%
  filter(CITY %in% city_filter) %>%
  ggplot(mapping = aes(x = Adjusted_Price)) +
  geom_histogram(bins = 50, fill = "blue", color = "black", alpha = 0.7) +
  facet_wrap(~ CITY, scales = "free") +
  scale_x_continuous()  # Optionally add breaks and labels here if needed
```

## Week 3 -


```{r}
# Possible way to get reset of lat/long
library(tidygeocoder)

lat_longs <- adjusted_data %>%
 geocode(full_adress, method = "osm", lat = Latitude, long = Longitude, full_results = TRUE)
```

